---
title: "Homework16"
format: html
editor: visual
---

The purpose of this file is to gain more practice applying the data analysis techniques detailed in GSWR Chapter 3 and functions inside the dplyr package

Set up

```{r}
#Clear the environment
rm(list = ls())
#Load in packages
library(tidyverse)
library(here)
```

Import and examine "Donnerville Percent Cover" practice data set

```{r}
#Import the data set
cover <- read.csv(here("data", "donnerville_percent_cover.csv"))
#Examine the data
glimpse(cover)
str(cover)
```

Rename some of the column headings

```{r}
cover <- cover %>%
  rename("Type_plot" = "Plot.type",
         "Dim_plot" = "Plot.dim",
         "Type_cover" = "cover_type")
```

Convert Site, Type_plot, and Type_cover into factor variables

```{r}
cover$Site <- as.factor(cover$Site)
cover$Type_plot <- as.factor(cover$Type_plot)
cover$Type_cover <- as.factor(cover$Type_cover)
```

Convert Date to a date variable

```{r}
cover$Date <- dmy(cover$Date)
```

Error check the data set and make corrections accordingly

```{r}
#Identify rows in Site that need repair for Donnerville 1
fix_S1 <- which(cover$Site %in% c("donnerville 1", "DONNERVILLE - 1"))
#Apply repair 
cover$Site[fix_S1] <- "Donnerville 1"

#Identify rows in Site that need repair for Donnerville 2
fix_S2 <- which(cover$Site == "donnerville 2")
#Apply repair 
cover$Site[fix_S2] <- "Donnerville 2"

#Identify rows in Site that need repair for Donnerville 3
fix_S3 <- which(cover$Site == "donnerville-3")
#Apply repair
cover$Site[fix_S3] <- "Donnerville 3"

#Identify rows in Type_cover that need repair for Bare
fix_TC1 <- which(cover$Type_cover == "bare")
#Apply repair
cover$Type_cover[fix_TC1] <- "Bare"

#Identify rows in Type_cover that need repair for Bryophytes
fix_TC2 <- which(cover$Type_cover == "bryophtes")
#Apply repair
cover$Type_cover[fix_TC2] <- "Bryophytes"

#Identify rows in Type_cover that need repair for Vascular
fix_TC3 <- which(cover$Type_cover == "vasc ")
#Apply repair
cover$Type_cover[fix_TC3] <- "Vascular"

#Check levels to ensure changes were effective
cover$Site <- factor(cover$Site)
levels(cover$Site)

cover$Type_cover <- factor(cover$Type_cover)
levels(cover$Type_cover)
```

Determine average percent cover by cover type for Donnerville 1

```{r}
filter(cover, Site == "Donnerville 1") %>%
  group_by(Type_cover) %>% 
    summarize(
      Mean_Percent = mean(Percent)
    )
```

Determine the average percent cover for vascular Type_cover

```{r}
filter(cover, Type_cover == "Vascular") %>%
  summarize(
    Mean_Percent = mean(Percent)
  )
```
