---
title: "Homework-15.qmd"
format: html
editor: visual
---

The purpose of this file is to provide a space to work through the material and examples in chapter 3 of GSWR

Set up

```{r}
#Clear the environment
rm(list = ls())
#Load packages
library(tidyverse)
library(here)
library(ggplot2)
library(dplyr)
```

Read in the dataset we will be working with

```{r}
compensation <- read.csv(here("data","compensation.csv"))
```

Examine the data and obtain summary statistics

```{r}
#Basic overview of the data frame
glimpse(compensation)
#Summary statistics
summary(compensation)
```

Access columns in the data frame using the select() function

```{r}
#Access the "Fruit" column
select(compensation, Fruit)
#Access every column except for the "Root" column
select(compensation, -Root)
```

Access rows in the data frame using the slice() function

```{r}
#Access the second row in the data frame
slice(compensation, 2)
#Access the second through tenth rows in the data frame
slice(compensation, 2:10)
#Access the second, third, and tenth rows of the data frame
slice(compensation, c(2,3,10))
```

Use the filter and with functions

```{r}
#Use with function to return boolean values
with(compensation, Fruit > 80)
#Use filter function to returns values satisfying a condition 
#Find spots in the data frame where fruit weighs more than 80 kgs
filter(compensation, Fruit > 80)
#Find spots in the data frame where fruit weighs more than 80 or less than 20 kgs
filter(compensation, Fruit > 80 | Fruit < 20)
```

Assign the values returned by a subset to an object for future use

```{r}
lo_hi_fruit <- filter(compensation, Fruit > 80, Fruit < 20)
```

Log transform the values stored in the Fruit column

```{r}
#Look at the first few rows before transformation
head(compensation)
#Perform the transformation
compensation <- mutate(compensation, logFruit = log(Fruit))
#Look at the first few rows after the transformation
head(compensation)
```

Sort the data by from lowest to highest (increasing order) values in the Fruit column

```{r}
arrange(compensation, Fruit)
```

Combine dplyr functions

```{r}
#Access Root values for which the associated Fruit values are greater than 80 kgs
select(filter(compensation, Fruit > 80), Root)
```

Access Root values for which the associated Fruit values are greater than 80 kgs using the pipe command

```{r}
compensation %>%
  filter(Fruit > 80) %>%
    select(Root)
```

Obtain the mean values for Fruit based on Grazing (Grazed or Ungrazed)

```{r}
mean.fruit <- summarize(
  group_by(compensation, Grazing), 
  meanFruit = mean(Fruit)
)
```

Obtain the mean values and standard deviations for Fruit based on Grazing (Grazed or Ungrazed) using the pipe command

```{r}
compensation %>%
  group_by(Grazing) %>%
  summarize(
    meanFruit = mean(Fruit),
    sdFruit = sd(Fruit)
  )
```

This looks great Jason - seems like you are humming along!
