---
title: "Homework 27"
format: html
editor: visual
---

Read through the code in this assignment. You will begin contributing to the code on line 36. Your tasks will be to:

-   Create a dataframe with just the complete cases

-   Review the 1-way ANOVA we ran during class with the full data set

-   Conduct a 2nd 1-way ANOVA using just the data set with complete cases and compare the results to the first analysis.

-   Conduct a linear regression to assess the relationship b/w Growth_rate_mm and Dried_mass_g, first with the complete data set and second with the complete cases data set and compare the results of the 2 analyses.

## Set up our workspace and read in the data

```{r}
rm(list = ls())
library(tidyverse)
library(ggfortify)
library(here)
```

Read in the data.

```{r}
df <- read.csv(here("Data/learn-about-NA.csv"))
str(df)
```

## Dealing with NAs in data analysis

Deciding how to treat NA values in your data can have critical impacts on your data analysis.

Let's look at an example.

Start by creating a data frame from df called df_All_NA, which includes the entire dataset including all NA values.

```{r}
df_All_NA <- df
```

Now use one of the methods we discussed in class on 13 November to create a second dataframe called `df_No_NA` that is just the complete cases of df.

```{r}
df_No_NA <- df[complete.cases(df), ] 
```

What if we wanted to test the hypothesis that there is no effect of fertilizer treatment on plant growth rate?

In class on Wednesday 13 November we ran a 1-way ANOVA on the complete data set. Let's review it here:

### First ANOVA using all the data

#### Plot and Guess

```{r}
ggplot(df_All_NA, aes(Treatment, Growth_rate_mm))+
  geom_boxplot() +
  geom_point() +
  theme_bw()
```

We decided that there was likely an effect of treatment on growth rate, but we weren't sure if both levels of the fertilizer treatment would differ from the control.

#### Create a model:

```{r}
grow.mod<-lm(Growth_rate_mm ~ Treatment, data = df_All_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod)
```

Looks pretty good.

#### Interpret results:

```{r}
anova(grow.mod)
```

Practice your interpretation.

Why are there 2 DF for the Treatment effect?

Answer: Because there are 3 levels of our Treatment variable: Control, Level 1, and Level 2. Number of groups - 1 = degrees of freedom therefore 3 - 1 = 2.

Why are there 46 DF for the Residuals? Remember that we determined that there were 49 values in the `Growth_rate_mm` variable using `sum(!is.na(variable))` like this:

```{r}
sum(!is.na(df$Growth_rate_mm))
```

Answer: There are 46 degrees of freedom because there are 49 datapoints in the Growth_rate_mm column after excluding NA values, and 3 parameters being estimated, 49 - 3 = 46 = degrees of freedom for the residuals.

So **there are only 49 data points used in the analysis**, even though the dataframe is 60 rows long!!!

```{r}
summary(grow.mod)
```

Practice your interpretation:

Does adding 1 dose of fertilizer change the growth rate significantly from the Control?

Answer: No, adding 1 dose of fertilizer does not significantly change the growth rate compared to the control. This is because the p-value = 0.0812, which is greater than 0.05.

Does adding 2 doses of fertilizer change the growth rate significantly from the Control?

Answer: Yes, adding 2 doses of fertilizer does significantly change the growth rate compared to the control. This is because the p-value = 0.0108, which is less than 0.05.

We won't worry about re-plotting our data right now.

### Second Attempt: Complete Cases

Your turn. Run the 1-way ANOVA again, this time with only the dataset that has complete cases and that we called `df_No_NA`

#### Plot and Guess

```{r}
ggplot(df_No_NA, aes(Treatment, Growth_rate_mm))+
  geom_boxplot() +
  geom_point() +
  theme_bw()
```

What do you think?

Answer: Examining the new plot, it seems the level 1 and level 2 treatments are more similar to one another. The means (center of the boxes) are closer to each other, and the medians represented by the black lines are almost perfectly aligned. Thus, it seems likely that both 1 and 2 doses of fertilizer will significantly differ from the control group.

#### Create a model:

```{r}
grow.mod.no.NA <- lm(Growth_rate_mm ~ Treatment, data = df_No_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod.no.NA)
```

What do you think? Same or different from our first 1-way ANOVA? Answer: Fairly similar to the 1-way ANOVA test autoplot graphs, thus our assumptions are okay to make.

#### Interpret results:

```{r}
anova(grow.mod.no.NA)
```

Practice your interpretation.

Why are there 2 DF for the Treatment effect? Answer: Because there are 3 levels of our Treatment variable: Control, Level 1, and Level 2. Number of groups - 1 = degrees of freedom therefore 3 - 1 = 2.

Why are there 33 DF for the Residuals? Answer: There are 33 degrees of freedom because there are 36 datapoints in the Growth_rate_mm column and 3 parameters being estimated, 36 - 3 = 33 = degrees of freedom for the residuals.

How would you write a results statement? Answer: There is no statistical significance between treatment group and growth rate, in other words, treatment group cannot be used as a predictor of growth rate (Degrees of Freedom = 2, F-value = 2.7906, p-value: 0.0759).

```{r}
summary(grow.mod.no.NA)
```

Practice your interpretation:

Does adding 1 dose of fertilizer change the growth rate significantly from the Control? Answer: No, adding 1 dose of fertilizer does not significantly change the growth rate compared to the control. This is because the p-value = 0.1139, which is greater than 0.05.

Does adding 2 doses of fertilizer change the growth rate significantly from the Control? Answer: Yes, adding 2 doses of fertilizer does significantly change the growth rate compared to the control. This is because the p-value = 0.0325, which is less than 0.05.

We won't worry about re-plotting our data right now.

## Linear Regression

Now conduct a linear regression to examine the relationship between `Growth_rate_mm` and `Dried_mass_g`, first with the complete data set and second with the complete cases data set and compare the results of the 2 analyses.

### First attempt: Full data set including NAs (use `df_All_NA`)

You know what to do here. Go for it.

#### Plot and guess

```{r}
ggplot(df_All_NA, aes(Dried_mass_g, Growth_rate_mm)) +
  geom_point() +
  theme_bw()
```

Based on this graph, it is difficult to discern, but there may be a slight positive correlation between dried mass and growth rate.

#### Create model

```{r}
mass.grow.mod <- lm(Growth_rate_mm ~ Dried_mass_g, data = df_All_NA)
```

#### Check assumptions

```{r}
autoplot(mass.grow.mod)
```

Don't worry if assumptions don't look perfect - focus on normality of the data. Based on the above, our assumptions are okay.

#### Interpret model

Use `anova()` first before using `summary()`

```{r}
anova(mass.grow.mod)
```

Why is there 1 degree of freedom for Growth_rate_mm and why are there 34 degrees of freedom for the Residuals? Answer: Since there are two variables we are investigating, growth rate and mass, we have Degrees of Freedom = \# Variables - 1, so 2 - 1 = 1. Similarly, for the residuals, since we have 2 variables and 36 data points after excluding rows congaing NAs, this gives us 36 - 2 = 34 degrees of freedom.

Based on the results of the ANOVA test, dried mass is a statistically significant predictor of growth rate (Degrees of Freedom = 1, F-value = 10.805, p-value = 0.002356).

```{r}
summary(mass.grow.mod)
```

Based on the results of this summary test, the relationship between growth rate and dried mass is statistically significant, with a p-value of 0.00236.

### Second attempt: Limited data set with only complete cases (use `df_No_NA`)

#### Plot and guess

```{r}
ggplot(df_No_NA, aes(Dried_mass_g, Growth_rate_mm)) +
  geom_point() +
  theme_bw()
```

Based on this graph, it is difficult to discern, but there may be a slight positive correlation between dried mass and growth rate.

#### Create model

```{r}
mass.grow.mod.no.NA <- lm(Growth_rate_mm ~ Dried_mass_g, data = df_No_NA)
```

#### Check assumptions

```{r}
autoplot(mass.grow.mod.no.NA)
```

Again, don't worry if the assumptions aren't perfect. How do they compare to the first analysis? Once again, given the above our assumptions are okay. The autoplot graphs generated appear almost identical to the previous autoplot graphs.

#### Interpret model

Use `anova()` first then `summary()`.

```{r}
anova(mass.grow.mod.no.NA)
```

Based on the results of the ANOVA test, dried mass is a statistically significant predictor of growth rate (Degrees of Freedom = 1, F-value = 10.805, p-value = 0.002356).

```{r}
summary(mass.grow.mod.no.NA)
```

Based on the results of this summary test, the relationship between growth rate and dried mass is statistically significant, with a p-value of 0.00236.

Answer the following questions:

1.  How many degrees of freedom are there for Growth_rate_mm? Why? Answer: One degree of freedom. Since there are two variables we are investigating, growth rate and mass, we have Degrees of Freedom = \# Variables - 1, so 2 - 1 = 1.

2.  How many degrees of freedom are there for Residuals? Why? Answer. Thirty four degrees of freedom. Since we have 2 variables and 36 data points, this gives us 36 - 2 = 34 degrees of freedom.

3.  How are the degrees of freedom in this second model different than the first? What does the answer tell you about the role of NA values in the regression analysis? The degrees of freedom are the same in the first in second model, which tells us that in the first model, the NA rows are ignored. However, it is still better to handle the NA values by removing them for cleaner analyses and conveying of results.
