---
title: "Homework18"
format: html
editor: visual
---

The purpose of this file is to gain more practice applying the data analysis techniques detailed in GSWR Chapter 4 and functions inside the dplyr package

```{r}
#Clear the environment 
rm(list = ls())
#Load in packages 
library(tidyverse)
library(here)
```

Read in the compensation data set

```{r}
compensation <- read.csv(here("Data","compensation.csv"))
```

Take a quick look at the data set

```{r}
glimpse(compensation)
```

Make a simple scatterplot of the data

```{r}
ggplot(compensation, aes(x=Root, y=Fruit)) +
  geom_point()
```

Change the "theme" of the scatterplot to override the default grey backround

```{r}
ggplot(compensation, aes(x=Root, y=Fruit)) +
  geom_point() +
  theme_bw()
```

Customize the size of the scatterplot points

```{r}
ggplot(compensation, aes(x=Root, y=Fruit)) +
  geom_point(size=5) +
  theme_bw()
```

Label the x and y-axes of the scatterplot

```{r}
ggplot(compensation, aes(x=Root, y=Fruit)) +
  geom_point(size=5) +
  xlab("Root Biomass") + 
  ylab("Fruit Production") +
  theme_bw()
```

Change the color of the points to correspond to the Grazing level

```{r}
ggplot(compensation, aes(x=Root, y=Fruit, color=Grazing)) +
  geom_point(size=5) +
  xlab("Root Biomass") +
  ylab("Fruit Production") +
  theme_bw()
```

Change the shape of the point to correspond to the Grazing level

```{r}
ggplot(compensation, aes(x=Root, y=Fruit, shape=Grazing)) +
  geom_point(size=5) +
  xlab("Root Biomass") +
  ylab("Fruit Production") +
  theme_bw()
```

Make a box-and-whisker plot of the data

```{r}
ggplot(compensation, aes(x=Grazing, y=Fruit)) +
  geom_boxplot() +
  xlab("Grazing treatment") +
  ylab("Fruit Production") +
  theme_bw()
```

Layer the raw data on top of the box-and-whisker plot with a few customizations

```{r}
ggplot(compensation, aes(x=Grazing, y=Fruit)) +
  geom_boxplot() + 
  geom_point(size=4, color='lightgray', alpha=0.5) +
  xlab("Grazing treatment") +
  ylab("Fruit Production") +
  theme_bw()
```

Answer the following questions: 1. Do plants with wider root diameters at the start of the experiment produce more fruit? -\> Yes 2. Do grazed or ungrazed plants produce more fruit -\> Grazed

Make a histogram of the data

```{r}
ggplot(compensation, aes(x=Fruit)) +
  geom_histogram()
```

Correct the histogram by adjusting the number of bins

```{r}
ggplot(compensation, aes(x=Fruit)) +
  geom_histogram(bins = 10)
```

Correct the histogram by adjusting the width of the bins

```{r}
ggplot(compensation, aes(x=Fruit)) + 
  geom_histogram(binwidth = 15)
```

Facet the histogram by the grazing level

```{r}
ggplot(compensation, aes(x=Fruit)) +
  geom_histogram(binwidth=15) + 
  facet_wrap(~Grazing)
```

Save the figure in the plots window to the working directory

```{r}
ggsave("ThatCoolHistogramAboutFruit.png")
```
