---
title: "Homework-28"
format: html
editor: visual
---

The purpose of this file is to provide a space to work through the material and examples in chapter 8 of GSWR

#### Set up
```{r}
#Clear the environment
rm(list=ls())
#Load in packaged
library(tidyverse)
library(here)
library(ggfortify)
library(gridExtra)
```

#### Load in and examine the dataset
```{r}
# Load in the dataset
compensation <- read.csv(here("Data","compensation.csv"))
# Examine the dataset
glimpse(compensation)
```

#### Construct basic plots

```{r}
# BASE scatterplot
base_scatter <- ggplot(data = compensation, aes(x = Root, y = Fruit)) +
  geom_point()
 # BASE box-and-whiskers plot
base_box <- ggplot(data = compensation, aes(x = Grazing, y = Fruit)) +
  geom_boxplot()

#Update the scatterplot to include a theme
base_scatter + theme_bw()
```

#### Place figures onto the same page
```{r}
grid.arrange(base_scatter, base_box, nrow = 1)
```

#### Modify the plot labels
```{r}
#Add axis labels
base_scatter + xlab("Root Biomass") + ylab("Fruit Production")
#Add a title 
base_scatter + ggtitle("Very necessary title")
#The following code accomplishes the same result as the above lines
base_scatter + labs(title = "Very necessary title",
                        x = "Root Biomass",
                        y = "Fruit Production")
```

#### Modify the axes' ranges
```{r}
base_scatter + xlim(0, 20) + ylim(0, 140)
```

#### Place text inside the figure
```{r}
base_scatter +
 annotate("text", x = c(6,8), y = c(105, 25),
 label = c("Text up here...","...and text down here"))
```

#### Use scale to edit tick placements
```{r}
base_scatter + scale_x_continuous(limits = c(4, 11), breaks = 4:11)
```

### Use scale to modify point colors to distinguish between groups
```{r}
ggplot(data = compensation, aes(x = Root, y = Fruit, colour = Grazing)) +
 geom_point() +
 scale_colour_manual(values = c(Grazed = "brown",
                                Ungrazed = "green"))
```

#### Perform a log transform
```{r}
base_box +
 scale_y_continuous(breaks = seq(from = 10, to = 150, by = 20),
                    trans = "log10")
```

#### Manuanlly adjust the scatterplot theme
```{r}
base_scatter +
 theme(
 panel.background = element_rect(fill = NA, colour = "black"),
 panel.grid.minor = element_blank(),
 panel.grid.major = element_line(colour = "lightblue")
 )
```

#### Adjust color and size of x-axis title and change the angle of the tick labels
```{r}
base_box +
 theme(
 axis.title.x = element_text(colour = "cornflowerblue",
 size =rel(2)),
 axis.text.x = element_text(angle = 45, size = 13, vjust = 0.5))
```

#### Alter level names
```{r}
base_box + scale_x_discrete(limits = c("Ungrazed", "Grazed"),
 labels = c("Control", "Grazed"))
```

#### Modify the key
```{r}
ggplot(compensation, aes(x = Root, y = Fruit, colour = Grazing)) +
 geom_point() +
 theme(legend.key = element_rect(fill = NA))
```

#### Remove the key
```{r}
ggplot(compensation, aes(x = Root, y = Fruit, colour = Grazing)) +
 geom_point() +
 theme(legend.position = "none")
```