---
title: "Biostats - Homework 20"
format: html
editor: visual
---

### Introduction

The objective of this assignment is to have you continue practicing the skills you have learned in Chapters 3 and 4 of Getting Started with R and reinforce the ideas we discussed in class regarding the Weissgerber et al. paper. Please complete all of the following steps:

1.  Write the code to set up your workspace.

```{r}
#Clear the environment 
rm(list = ls())
#Load in the packages 
library(tidyverse)
library(here)

```

2.  Read in the dataset called soil_data.csv and name it "soil."

```{r}
soil <- read.csv(here("Data", "soil_data.csv"))
```

3.  Create a histogram for the SOM_percent column in the dataframe. Accept the default number of bins.

```{r}
ggplot(soil, aes(SOM_percent)) +
  geom_histogram()
```

How would you describe the mean_SOM data based on viewing the histogram? Write your answer below:
The data follows a right skewed distribution, most SOM in the sample seems to fall below 15%

4.  Create a new dataframe called SOM by calculating for each forest the mean soil organic matter and the standard deviation of soil organic matter. Make sure to name variables you create "mean_SOM" and "sd_SOM" so that the next code chunk I give you works.

```{r}
SOM <- soil |>
  group_by(Forest) |>
    summarize(
      mean_SOM = mean(SOM_percent),
      sd_SOM = sd(SOM_percent)
    )
```

Now summarize Nitrogen
```{r}
Nit <- soil |> 
    group_by(Forest) |> 
      summarize(
        mean_n = mean(Nitrate_N_ppm, na.rm = T),
        sd_n = sd(Nitrate_N_ppm, na.rm = T)
      ) |> view()
```

5.  Now run the following code chunk which will create a bar plot of mean soil organic matter per forest, with standard deviation as error bars.

```{r}
ggplot(SOM, aes(Forest, mean_SOM))+
  geom_errorbar(aes(ymax = mean_SOM + sd_SOM, ymin = mean_SOM - sd_SOM), width = 0.2)+
  geom_col(fill = "white", color = "black")+
  ylab("Mean percent soil organic matter")+
  xlab("Forest name")
```

Based on what you see in the bar chart, how would you describe the pattern of SOM between the 4 different forests? Write your answer here:
On average, the KIP Tract forest has the greatest percent SOM and largest variation, followed by Donnerville, then South Hammond, then Degrasse (based on bar heights). While not as obvious as on the histogram, if we imagine a line around 15, we can see the majority of the data fall below this. The bar chart also reveals that much of what's causing the "right-skewness" is due to the KIP tract as it reaches almost 40 in its variation, which is considerably higher than the others. 

Now create a boxplot of the SOM_percent data. Use Forest as the x-axis variable and overlay the point data on top of the boxplot. Make the points blue and somewhat transparent. Call the x-axis "Forest name" and the y-axis "Percent soil organic matter" (hint: See homework 18...).

```{r}
ggplot(soil, aes(x = Forest, y = SOM_percent)) +
  geom_boxplot() +
  geom_point(size=4, color='blue', alpha=0.2)
```

Based on what you see in the box plot, write out your answers the following questions:

-   how would you describe the pattern of SOM between the 4 different forests?
    The Kip Tract forest has the greatest average SOM_percent as indicated by the black bar in the box being higher than the others, followed by Donnerville, then South Hammond, then Degrasse. We also see that the KIP tract has the largest interquartile ranges (white space of the box above and below the black line), while Donnerville and South Hammond are similar, and Degrasse has small interquartile ranges. 

-   What do you notice that is different from how the data were displayed in the bar plot?
  With the bar plot, a line is created to indicate the average for a forest group and the bar simply starts from 0 and extends to the line to indicate it. With the box plot, a line is created at the average for a forest group, but the bar starts there and extends to the values of the interquartile ranges (so it no longer has to touch 0).
-   Which plot type is more informative? Explain your answer.
The box plot is more informative since it includes interquartile ranges while the bar plot does not. A box plot also allows the variance under the average to be seen while the bar plot does not.
