---
title: "Homework-25"
format: html
editor: visual
---

This file was constructed for the purpose of working through the examples in chapters 5.6-5.7 of the textbook (Getting started with R)

Set up

```{r}
#Clear the environment
rm(list = ls())
#Load in packages
library(tidyverse)
library(here)
library(ggfortify)
```

Load in and examine the dataset
```{r}
#Load in the dataset
daphnia <- read.csv(here("Data", "Daphniagrowth.csv"))
#Examine the dataset
glimpse(daphnia)
```

Convert applicable columns to factor variable types
```{r}
daphnia$parasite = as.factor(daphnia$parasite)
```

Construct a box-and-whisker plot to get an idea of the relationship we are interested in investigating, the effect parasites have on the growth rate of daphnia
```{r}
ggplot(daphnia, aes(x=parasite, y=growth.rate)) +
  geom_boxplot() +
  theme_bw()
```

We can also swap the x- and y- axes using coord_flip
```{r}
ggplot(daphnia, aes(x=parasite, y=growth.rate)) +
  geom_boxplot() +
  theme_bw() +
  coord_flip()
```

Conduct an ANOVA test
```{r}
#Fit the model
model_grow <- lm(growth.rate ~ parasite, data=daphnia)
#Assess assumptions
autoplot(model_grow, smooth.color = NA)
#Run the anova test
anova(model_grow)
```

View contrasts using summary table 
```{r}
summary(model_grow)
```

Treatment contrasts can also be calculated manually by obtaining and viewing means
```{r}
#Calculate means
sumDat <- daphnia |>
  group_by(parasite) |>
    summarize(
      meanGR = mean(growth.rate)
    )
#View means 
sumDat
```

Construct a graph that models growth rate among different treatment levels and also include the means
```{r}
ggplot(daphnia, aes(x=growth.rate, y=parasite, color=parasite)) +
  geom_point(size=3, alpha=0.5) +
  geom_point(data=sumDat, aes(x=meanGR, y=parasite, color= parasite), shape=18, size=7)
```